# Reproducible Research: Peer Assessment 1
NOTE: Set working directory to directory containing this file.

Load required libraries
```{r}
library(ggplot2)
library(data.table)
```

## Loading and preprocessing the data
The data is loaded to a 'data table', directly from the zipped file:
```{r}
step.data <- data.table(read.csv(unz("activity.zip", "activity.csv")))
setkey(step.data, date, interval)
```
A quick check of the data is performed as follows:
```{r}
head(step.data)
dim(step.data)
```
This is consistent with expectations.

## What is mean total number of steps taken per day?
Data tables can easilly be aggregated by key values.  Above, I set 'date' to be a key and then calculated the sum aggregated by this key.  From this, the ggplot library is used to plot a histogram.  Then, the mean and median total number of steps are calculated and reported.

```{r}
# aggregate total steps per day
sum.steps.per.day <- step.data[, list(sumSteps = sum(steps)), by = date]

# plot histogram of total steps per day
g1 <- ggplot(sum.steps.per.day, aes(x = sumSteps))
g1 <- g1 + geom_histogram(binwidth = 5000)
g1 <- g1 + xlab("Number of Steps in a Day")
g1 <- g1 + ylab("Frequency")
g1 <- g1 + ggtitle("Histogram of Total Number of Steps in a Day")
g1 <- g1 + xlim(0, 25000)
print(g1)

cat(sprintf("Mean Number of Total Steps per Day: %.1f", 
                        sum.steps.per.day[, mean(sumSteps, na.rm = T)]))
cat(sprintf("Median Number of Total Steps per Day: %.1f", 
                        sum.steps.per.day[, median(sumSteps, na.rm = T)]))
```

## What is the average daily activity pattern?

```{r}
ave.steps.per.int <- 
        step.data[, list(aveSteps = mean(steps, na.rm = T)), by = interval]
g2 <- ggplot(ave.steps.per.int, aes(y = aveSteps, x = interval))
g2 <- g2 + geom_line()
g2 <- g2 + xlab("Interval Identifier")
g2 <- g2 + ylab("Mean Number of Steps")
g2 <- g2 + ggtitle("Mean Number of Steps versus Interval Identifier")
print(g2)

cat(sprintf(
        "The inteval with the maximum number of steps, on average, is: %i", 
        ave.steps.per.int[, interval[which.max(aveSteps)]]))
```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
