# Reproducible Research: Peer Assessment 1
NOTE: Set working directory to directory containing this file.

Load required libraries
```{r}
library(ggplot2)
```

## Loading and preprocessing the data
The data is loaded, directly from the zipped file:
```{r}
step.data <- read.csv(unz("activity.zip", "activity.csv"))
```
A quick check of the data is performed as follows:
```{r}
head(step.data)
dim(step.data)
```
This is consistent with expectations.

## What is mean total number of steps taken per day?
First, an aggregate sum of the steps, split by day, is calculated.  From this, the ggplot library is used to plot a histogram.  Then, the mean and median total number of steps are calculated.

```{r}
# aggregate total steps per day
steps.per.day <- aggregate(steps ~ date, data = step.data, sum)

# plot histogram of total steps per day
g <- ggplot(steps.per.day, aes(x = steps))
g <- g + geom_histogram(binwidth = 5000)
g <- g + xlab("Number of Steps in a Day")
g <- g + ylab("Frequency")
g <- g + ggtitle("Histogram of Total Number of Steps in a Day")
g <- g + xlim(0, 25000)
print(g)

cat(sprintf("Mean Number of Total Steps per Day: %.1f", 
                                                mean(steps.per.day$steps)))
cat(sprintf("Median Number of Total Steps per Day: %.1f", 
                                                median(steps.per.day$steps)))
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
