---
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
# Reproducible Research: Peer Assessment 1

Load required libraries
```{r}
library(ggplot2)
library(data.table)
library(grid)
library(gridExtra)
```

## Loading and preprocessing the data
In order to run this document, set working directory to directory containing this file.  The data is loaded to a 'data table', directly from the zipped file:
```{r}
step.data <- data.table(read.csv(unz("activity.zip", "activity.csv")))
setkey(step.data, date, interval)
```
A quick check of the data is performed as follows:
```{r}
head(step.data)
dim(step.data)
```
This is consistent with expectations.

## What is mean total number of steps taken per day?
Data tables can easily be aggregated by key values.  Above, I set 'date' to be a key and then calculated the sum aggregated by this key.  From this, the ggplot library is used to plot a histogram.  Then, the mean and median total number of steps are calculated and reported.  Note that missing values are to be ignored so I set `na.rm=T` in the code below.
```{r}
# aggregate total steps per day
sum.steps.per.day <- step.data[, list(sumSteps = sum(steps, na.rm = T)), by = date]

# plot histogram of total steps per day
g1 <- ggplot(sum.steps.per.day, aes(x = sumSteps))
g1 <- g1 + geom_histogram(binwidth = 5000)
g1 <- g1 + xlab("Number of Steps in a Day")
g1 <- g1 + ylab("Frequency")
g1 <- g1 + ggtitle("Histogram of Total Number of Steps in a Day")
g1 <- g1 + xlim(0, 25000)
print(g1)

# write out results
cat(sprintf("Mean Number of Total Steps per Day: %.1f", 
                        sum.steps.per.day[, mean(sumSteps, na.rm = T)]))
cat(sprintf("Median Number of Total Steps per Day: %.1f", 
                        sum.steps.per.day[, median(sumSteps, na.rm = T)]))
```

## What is the average daily activity pattern?
Using a similar approach to above, the data table aggregate mean is computed by interval.  
```{r}
# aggregate mean steps by interval
ave.steps.per.int <- 
        step.data[, list(aveSteps = mean(steps, na.rm = T)), by = interval]

# plot time series
g2 <- ggplot(ave.steps.per.int, aes(y = aveSteps, x = interval))
g2 <- g2 + geom_line()
g2 <- g2 + xlab("Interval Identifier")
g2 <- g2 + ylab("Mean Number of Steps")
g2 <- g2 + ggtitle("Mean Number of Steps versus Interval Identifier")
print(g2)

# write out results
cat(sprintf(
        "The inteval with the maximum number of steps, on average, is: %i", 
        ave.steps.per.int[, interval[which.max(aveSteps)]]))
```

## Inputing missing values
The number of missing data is computed as follows:
```{r}
cat(sprintf(
        "The total number of missing data is: %i", 
        step.data[, sum(is.na(steps))]))
```

Here, I will 'correct' the missing data by substituting in the mean steps for that interval across all of the days of collected data:
```{r}
# correction to data
step.data[, stepsCor := 
                ifelse(
                        is.na(steps), 
                        ave.steps.per.int[interval == interval, aveSteps], 
                        steps
                )
]

# aggregate data
sum.steps.per.day <- step.data[, 
                list(sumSteps = sum(steps, na.rm = T), sumStepsCor = sum(stepsCor)), 
                by = date]
print(sum.steps.per.day)

# plot data
g3 <- ggplot(sum.steps.per.day, aes(x = sumSteps))
g3 <- g3 + geom_histogram(binwidth = 5000)
g3 <- g3 + xlab("Number of Steps in a Day")
g3 <- g3 + ylab("Frequency")
g3 <- g3 + ggtitle("Missing Data Not Corrected")
g3 <- g3 + xlim(0, 25000) + ylim(0, 40)  

g4 <- ggplot(sum.steps.per.day, aes(x = sumStepsCor))
g4 <- g4 + geom_histogram(binwidth = 5000)
g4 <- g4 + xlab("Number of Steps in a Day")
g4 <- g4 + ylab("Frequency")
g4 <- g4 + ggtitle("Missing Data Corrected")
g4 <- g4 + xlim(0, 25000) + ylim(0, 40)

g4 <- arrangeGrob(g3, g4, ncol = 1, 
                main = "Histograms of Total Number of Steps in a Day")
print(g4)

# write out results
summary(sum.steps.per.day)
```
As can be observed in the summary table, correcting the missing data has increased the mean and median.  The reason is that if the missing values are simply ignored, the sums of 'missing values' will be biased towards zero.  Hence, the data is biased which is manifested as a right skewed dikstribution.  By correcting the missing data, we correct for this bias.

## Are there differences in activity patterns between weekdays and weekends?
The following code identifies whether days are weekday or weekends.  Subsequently, a  plot is generated for the mean number of steps versus the interval, faceted by whether the day is during the week or at the weekend. 
```{r}
# create factor to identify whether day is week or weekend
step.data[, day.type := 
                        ifelse(weekdays(as.Date(date),T) %in% 
                        c("Sat", "Sun"), "week.end.day", "week.day")]

# aggregate data 
ave.steps.per.int <- 
                        step.data[, list(aveSteps = mean(stepsCor)), 
                        by = c("interval", "day.type")]

# plot data
g5 <- ggplot(data = ave.steps.per.int, aes(x = interval, y = aveSteps))
g5 <- g5 + geom_line() 
g5 <- g5 + facet_grid(day.type ~ .)
g5 <- g5 + xlab("Interval Identifier")
g5 <- g5 + ylab("Mean Number of Steps")
g5 <- g5 + ggtitle("Mean Number of Steps versus Interval Identifier")
print(g5)
```